Constraints on the mass and occupation fraction of dark matter halos hosting Lyman-alpha emitters at $z\sim 3$

### 2013-06-27 paper/ms.tex
minor modifications: first draft to be sent to friends

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$, by comparing the number density and the angular
correlation function between mock and observed fields. The mock fields
are constructed in a large cosmological N-body simulation using a
model where a dark matter halo can only host one LAE with a
probability $f_{\rm occ}$ if its mass is within the mass range
delimited by two threshold values  $M_{\rm min}$ and $M_{\rm max}$. We
conclude that the number density and spatial clustering information
available from observations is not sufficient to derive a unique and narrow set of
values for the parameters in the model. We find three different
families of models based on their mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ dex and occupation
fraction $f_{\rm occ}$.  However, there is a dominant familiy composed
by models with a narrow mass range $\Delta M<1.0$ dex, a low
occupation fraction $f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm
  max}< 10^{12}$\hMsun.  This finding suggests that the most massive
dark matter halos at that epoch do not host the brightest LAEs. This
also gives support to observational evidence concluding that only a
small fraction of star forming galaxies can be actually detected as
LAEs. We provide tables and mock catalogs for the successful models. 
\end{abstract}

@@

\begin{equation}
\xi(\theta) = \left(\frac{\theta}{\theta_{0}}\right)^{-\gamma}
\label{eq:fitting}
\end{equation}

@@

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the preferred
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method after applying the observational constraints 
in the occupation fraction and the ACF. Here we consider a model
consistent if there is an overlap of $1-sigma$ in its values for the
correlation length $\theta_0$ and the power $\gamma$ in the
correlation function.

@@

Matching the galaxy surface number density statistics sets the median
mass of all successful models in the range $10^{10}-10^{12}$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 90000 possible combination of parameters.

@@

Considering the additional evidence for a low escape fraction we can
say that the preferred models are in families (1) and (2) with a
clear majority composed by those with narrow mass range and low occupation
fraction.

Using abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} report that the instantaneous star
formation efficiency (star formation rate divided by the stellar mass)
presents a clear maximum around $10^{11.7}$\Msun \ at all redshifts
$z<4$.

This mass scale is strictly superior to the great majority of  $M_{\rm max}$
values allowed in the models with low escape fraction. Using the results published
in \cite{Behroozi2013a} we find that the typical stellar mass in halos
of $10^{11.4}$\hMsun \ is $(1.0\pm0.3)\times 10^{9.0}$ \hMsun, while
their star formation rate is in the range $0.6\pm 0.2$ Msun yr$^{-1}$.
This halo mass range around $10^{11.4}$ \hMsun \ is the lower bound of
what is computed in the abundance matching model in order to fit the
observational data for the stellar mass function based on the
observations of Lyman Break Galaxies.

All the preferred models have a halo mass range with $M_{\rm
  min}<10^{11.4}$\hMsun. This suggests that a detailed and careful
study of the spectral and photometric properties of LAEs coupled to the kind of
analysis performed in this paper can be a guide in the study of the
properties of low mass dark matter halos at $z=3.1$, extending the
capabilities of abundance matching methods.

>>>>

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$, by comparing the number density and the angular
correlation function between mock and observed fields. The mock fields
are constructed in a large cosmological N-body simulation using a
model where a dark matter halo can only host one LAE with a
probability $f_{\rm occ}$ if its mass is within the mass range
delimited by two threshold values  $M_{\rm min}$ and $M_{\rm max}$. We
find three different families of parameters based on their mass range
$\Delta M \equiv \log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ dex and occupation
fraction $f_{\rm occ}$.  From this we  conclude that the number
density and spatial clustering information available from observations
is not sufficient to derive a unique and narrow set of values for the
parameters in the model.  However, there is a dominant familiy composed
by models with a narrow mass range $\Delta M<1.0$ dex, a low
occupation fraction $f_{\rm occ}\leq 0.3$ and a maximum mass $M_{\rm
  max}< 10^{12}$\hMsun.  This suggests that the most massive
dark matter halos at that epoch do not host the brightest LAEs, also
giving support to some observational evidence indicating that only a small
fraction of star forming galaxies can be actually detected as LAEs. We
provide tables and mock catalogs for the successful models.  
\end{abstract}

@@

\begin{equation}
\xi(\theta) = \left(\frac{\theta}{\theta_{0}}\right)^{\gamma}
\label{eq:fitting}
\end{equation}

@@

In Figure \ref{fig:restriction_mock_and_f_occ_corr} we present the preferred
models in the planes $M_{\rm min}-M_{\rm  max}$ and $M_{\rm min}-f_{\rm occ}$
for the {\tt match} method after applying the observational constraints 
in the occupation fraction and the ACF. Here we consider a model
consistent if there is an overlap of $1-\sigma$ in its values for the
correlation length $\theta_0$ and the power $\gamma$ in the
correlation function.

@@

Matching the galaxy surface number density statistics sets the median
mass of all successful models in the range $(10^{10}-10^{12})$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 90000 possible combination of parameters.

@@

Considering the additional evidence for a low escape fraction we can
say that the preferred models are in families (1) and (2) with a
clear majority composed by those with narrow mass range.

Abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} work based on observational
results for Lyman Break Galaxies. In thos studies the minimum halo
mass considered to be relevant in their analysis is
$10^{11.4}$\hMsun. For that mass, \cite{Behroozi2013a} report a
stellar mass mass around $(1.0\pm0.3)\times 10^{9.0}$
\hMsun, while their star formation rate is in the range $0.6\pm 0.2$
Msun yr$^{-1}$, which nevertheless are close to the lower bound of
values inferred for LAEs at high redshift
\citep{Gawiser2007,Nilsson2009,Pentericci2009}.

In our results, all the preferred models have a halo mass range lower
than the minimum of $M_{\rm min}<10^{11.4}$\hMsun considerend in
abundance matching at $z=3$. This mass scale is also strictly superior to
the great majority of  $M_{\rm max}$ values allowed in the models with
low escape fraction.   This suggests that a detailed and careful study
of the s ectral and photometric properties of LAEs coupled to the kind
of analysis performed in this paper can be a guide in the study of the
properties of low mass dark matter halos at $z=3.1$, extending the
capabilities of abundance matching methods.
